Validating risk scores - Czech data

### 2021-01-25 other_outputs/risk_scores_consortium.Rmd
Another update to risk scores

data <- read_data_for_analysis()
```

```{r}
first_non_NA <- function(x) {
  if(length(x) > 2) {
    stop("Wrong length")
  } else if(length(x) == 1) {
    x
  } else {
    if(is.na(x[1])) {
      x[2]
    } else {
      x[1]
    }
  }

}

@@

@@

```{r}
data_risk <- data$patient_data %>%
  mutate(mortality  = case_when(
    last_record >= 30 ~ "alive",
    outcome == "Discharged" ~ "alive",
    outcome == "Death" ~ "dead",
    TRUE ~ "alive")) %>%
  inner_join(marker_data_admission, by = c("patient_id", "hospital_id")) %>%
  mutate(creatinin = if_else(is.na(creatinin.x), creatinin.y, creatinin.x),
         pt_inr = if_else(is.na(pt_inr.x), pt_inr.y, pt_inr.x),
         albumin = if_else(is.na(albumin.x), albumin.y, albumin.x),
         ) %>%
  select(-creatinin.x, -creatinin.y, -pt_inr.x, -pt_inr.y, -albumin.x, -albumin.y) %>%
  #avoid including some derived quantities
  select(-age_norm, -high_creatinin, -high_pt_inr, -low_albumin, -heart_problems, -lung_problems, -markers_out_of_range, -antibiotics, -macrolides, -comorbidities_sum_na) %>%
  #avoid columns that are not useful
  select(-hospital_id, -patient_id, -pcr_positive, -FiO2, -urea,
          -discontinued_medication, -last_record, -outcome,  
         #-best_supportive_care_from,
         -starts_with("ever_"), ever_hcq,
         -starts_with("any_"), 
         -starts_with("worst"), worst_breathing_s,
         -had_invasive,
         -first_day_invasive, -last_day_invasive,
      #   -startsWith("took"),
         -took_hcq,
         -took_convalescent_plasma, -obesity, -took_favipiravir, -took_az,
         -took_dexamethasone, -took_remdesivir, -IgG, -Horowitz_index,
         -has_hypertension_drugs
         ) %>%
  #convert to factors to use with mice
  mutate(sex = factor(sex), 
         breathing_s = factor(breathing_s, ordered = FALSE), 
         worst_breathing_s = factor(worst_breathing_s, ordered = FALSE), 
         requires_oxygen = breathing_s != "AA",
         across(all_of(names(.)[sapply(., typeof) == "logical"]), factor))
```

@@

```{r}
where_impute <- is.na(data_risk)
columns_to_impute <- c("lactate_dehydrogenase", 
                       "lymphocyte_count",
                       "SpO2",
                       "SpO2_native",
                       "CRP",
                       "d_dimer",
                       "COPD",
                       "other_lung_disease",
                       "has_hypertension_drugs",
                       "ischemic_heart_disease",
                       "heart_failure")

@@

predictor_matrix[, "requires_oxygen"] <- 0
predictor_matrix[, "mortality"] <- 0
predictor_matrix["oxygen_flow", "worst_breathing_s"] <- 0
low_obs_count <- c("lactate_dehydrogenase", "AST", "ferritin", "IL_6", "leukocyte_count", "PEEP")
predictor_matrix[low_obs_count, ] <- 0
predictor_matrix[low_obs_count, c("age","sex")] <- 1

@@

@@

```{r}
age_baseline <- function(data) {
  data %>% mutate(age_norm = (age - 65) / 20 - 2) %>%
    pull(age_norm)
}
```

>>>>

data <- read_data_for_analysis()

columns_to_impute <- c("lactate_dehydrogenase", 
                       "lymphocyte_count",
                       "SpO2",
                       "SpO2_native",
                       "CRP",
                       "d_dimer",
                       "COPD",
                       "other_lung_disease",
                       "has_hypertension_drugs",
                       "ischemic_heart_disease",
                       "heart_failure")
```

```{r}
first_non_NA <- function(x) {
  if(all(is.na(x))) {
    NA
  } else {
    ind <- which(!is.na(x))[1]
    x[ind]
  }
}

@@

marker_data_2_5_to_impute <- data$marker_data_wide %>%
  select(hospital_id, patient_id,  day, any_of(columns_to_impute)) %>%
  filter(day >= 2 & day <= 5) %>%
  arrange(hospital_id, patient_id,  day) %>%
  group_by(patient_id, hospital_id) %>%
  summarise(across(everything(), first_non_NA), .groups = "drop") %>%
  select(-day)

columns_2_5 <- setdiff(names(marker_data_2_5_to_impute), c("day", "hospital_id", "patient_id"))

@@

```{r}
data_risk <- data$patient_data %>%
  mutate(mortality  = case_when(
    last_record >= 30 ~ "alive",
    outcome == "Discharged" ~ "alive",
    outcome == "Death" ~ "dead",
    TRUE ~ "alive")) %>%
  inner_join(marker_data_admission, by = c("patient_id", "hospital_id")) %>%
  mutate(creatinin = if_else(is.na(creatinin.x), creatinin.y, creatinin.x),
         pt_inr = if_else(is.na(pt_inr.x), pt_inr.y, pt_inr.x),
         albumin = if_else(is.na(albumin.x), albumin.y, albumin.x),
         ) %>%
  select(-creatinin.x, -creatinin.y, -pt_inr.x, -pt_inr.y, -albumin.x, -albumin.y)

# Add data 2 to 5
data_risk <- data_risk %>%
  inner_join(marker_data_2_5_to_impute, by = c("patient_id", "hospital_id"),
           suffix = c("", "_2_5"))

  #avoid including some derived quantities
data_risk <- data_risk %>%
  select(-age_norm, -high_creatinin, -high_pt_inr, -low_albumin, -heart_problems, -lung_problems, -markers_out_of_range, -antibiotics, -macrolides, -comorbidities_sum_na) %>%
  #avoid columns that are not useful
  select(-hospital_id, -patient_id, -pcr_positive, -FiO2, -urea,
          -discontinued_medication, -outcome,  
         #-best_supportive_care_from,
         -starts_with("ever_"), ever_hcq,
         -starts_with("any_"), 
         -starts_with("worst"), worst_breathing_s,
         -had_invasive,
         -first_day_invasive, -last_day_invasive,
         -first_wave,
      #   -startsWith("took"),
         -took_hcq,
         -took_convalescent_plasma, -obesity, -took_favipiravir, -took_az,
         -took_dexamethasone, -took_remdesivir, -IgG, -Horowitz_index,
         -has_hypertension_drugs
         ) %>%
  #convert to factors to use with mice
  mutate(sex = factor(sex), 
         breathing_s = factor(breathing_s, ordered = FALSE), 
         worst_breathing_s = factor(worst_breathing_s, ordered = FALSE), 
         requires_oxygen = breathing_s != "AA",
         across(all_of(names(.)[sapply(., typeof) == "logical"]), factor))

# Nelson-Aalen for outcome
data_risk <- data_risk %>% 
  mutate(mortality_l = mortality == "dead") %>%
  mutate(
         nelsonaalen = mice::nelsonaalen(., timevar = "last_record", statusvar = "mortality_l")) %>%
  select(-last_record, -mortality_l)
```

@@

```{r}
where_impute <- is.na(data_risk)

@@

predictor_matrix[, "requires_oxygen"] <- 0
predictor_matrix[, "mortality"] <- 0
predictor_matrix["oxygen_flow", "worst_breathing_s"] <- 0
low_obs_count <- c("lactate_dehydrogenase", "AST", "ferritin", "IL_6", "leukocyte_count", "PEEP")
predictor_matrix[low_obs_count, ] <- 0
predictor_matrix[low_obs_count, c("age","sex")] <- 1
predictor_matrix["lactate_dehydrogenase", "sex"] <- 0

@@

no_nelsonaalen <- c("best_supportive_care_from")
predictor_matrix[no_nelsonaalen, "nelsonaalen"] <- 0

# Use future values only for the corresponding predictors
for(c in columns_2_5) {
  c_2_5 <- paste0(c, "_2_5")
  # The same predictor as for the original column
  predictor_matrix[c_2_5, ] <- predictor_matrix[c, ] 
  predictor_matrix[, c_2_5] <- 0

  predictor_matrix[c, paste0(c, "_2_5")] <- 1 
}

@@

```{r}
age_baseline <- function(data) {
  # Comparison of the characteristics, morbidity, and mortality of COVID-19 and seasonal influenza: a nationwide, population-based retrospective cohort study
  # DOI: 10.1016/S2213-2600(20)30527-0
  # https://www.thelancet.com/journals/lanres/article/PIIS2213-2600(2030527-0/fulltext#sec1
  # Supplementary Table 3
  data %>% mutate(ifr_perc = case_when(
    age < 18 ~ 0.7,
    age <= 30 ~ 0.5,
    age <= 40 ~ 1.4,
    age <= 50 ~ 2.7,
    age <= 60 ~ 6.1,
    age <= 70 ~ 13.7,
    age <= 80 ~ 23.1,
    age <= 90 ~ 32.4,
    TRUE ~ 36.3),
    ifr = ifr_perc / 100,
    lp = log(ifr) - log1p(-ifr)) %>%
    pull(lp) 
}
```
